---
title: "Framewise displacement across days and runs"
author: "Benjamin Meyer"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 8, fig.height = 5.5, dpi = 200,
  dev.args = list(bg = "white")
)
library(ggplot2)
library(tidyr)
library(dplyr)
library(rstatix)
library(knitr)

# Plotting helpers shared with the tSNR report.
source("qc_plot_helpers.R")
theme_set(qc_theme())

# PNG copies of the figures, for the GitHub-readable summary (RESULTS.md).
dir.create("figures", showWarnings = FALSE)
save_fig <- function(name, plot, w = 17, h = 12) {
  ggsave(paste0(name, ".pdf"), plot, width = w, height = h, units = "cm")
  ggsave(file.path("figures", paste0(name, ".png")), plot,
         width = w, height = h, units = "cm", dpi = 200)
}
```

This is the tSNR analysis repeated on head motion, with the echo factor removed:
framewise displacement is estimated once per BOLD series, so there is one value
per subject and run rather than one per echo. Two dependent variables are carried
through in parallel, **mean FD** and **max FD**, both in mm.

## Read the data

`fd_results.csv` is the long-format output of `extract_fd.py`: one row per
subject and run, with the acquisition day taken from the `ses-` entity of the
confounds filename.

```{r read}
fd <- read.csv("fd_results.csv") |>
  mutate(ID = factor(ID), run = as.integer(run), day = as.integer(day))

str(fd)
```

```{r coverage}
fd |>
  group_by(run, day) |>
  summarise(
    n_subjects  = dplyr::n(),
    n_vols      = paste0(min(n_vols), if (min(n_vols) != max(n_vols)) paste0("-", max(n_vols)) else ""),
    .groups = "drop"
  )
```

Runs differ slightly in length — 325 volumes on Days 1 and 3, 333 on Day 2. Mean
FD is an average and so unaffected; max FD is an extreme value and could in
principle be biased upwards by a longer run, but the difference is 2.5% of the
series and max FD is essentially uncorrelated with run length across the dataset
(Spearman *r* = `r round(cor(fd$max_FD, fd$n_vols, method = "spearman"), 3)`), so
it is not a material confound.

## Run-to-day mapping

Unchanged from the tSNR report (`eeg_fmri_ext_SNR_1_v2.m:63`,
`rest_day = [1 2 2 2 2 3]`; runs 2-5 carry EEG):

| Run | Session | EEG |
|-----|---------|-----|
| 1 | `ses-Day1_task-rest1` | no |
| 2 | `ses-Day2_task-rest2` | **yes** |
| 3 | `ses-Day2_task-rest3` | **yes** |
| 4 | `ses-Day2_task-rest4` | **yes** |
| 5 | `ses-Day2_task-rest5` | **yes** |
| 6 | `ses-Day3_task-rest6` | no |

The same two questions as before: **across days** (runs 1, 2, 6 — one run per
day) and **within Day 2** (runs 2-5 — the four EEG runs).

## Choice of test

The tSNR report used a repeated-measures ANOVA, and that is kept here as the
primary test so the two reports are comparable. FD, however, is a strictly
positive, right-skewed quantity — max FD especially, since it is the largest of
~330 values — so the normality assumption needs checking rather than assuming.
Each analysis below therefore reports, alongside the ANOVA:

- **Shapiro-Wilk on the paired differences**, which is the assumption the
  repeated-measures ANOVA and paired *t*-tests actually rest on;
- **Friedman's test with Kendall's *W***, the rank-based equivalent of the
  one-way within-subject ANOVA (the "non-parametric test?" noted in
  `instructions.md`);
- **Wilcoxon signed-rank post-hocs** next to the paired *t*-tests.

Where the two disagree, the non-parametric result is the one to trust.

```{r analysis-fns}
# Prepare one analysis set: pick the runs, label the grouping factor, keep only
# subjects with a value in every cell (the design is fully within-subject).
prep_fd <- function(dat, dv, runs, labels) {
  out <- dat |>
    filter(run %in% runs) |>
    mutate(grp = factor(run, levels = runs, labels = labels),
           y   = .data[[dv]])
  complete <- out |>
    group_by(ID) |>
    filter(dplyr::n() == length(runs)) |>
    ungroup() |>
    mutate(ID = droplevels(ID))
  attr(complete, "n_all") <- n_distinct(out$ID)
  add_cells(complete)
}

# Every test for one analysis set.
fd_fit <- function(d) {
  wide <- d |>
    select(ID, grp, y) |>
    pivot_wider(names_from = grp, values_from = y)
  ref  <- wide[[2]]
  diffs <- lapply(wide[-(1:2)], function(col) col - ref)
  shapiro <- data.frame(
    comparison = paste(names(wide)[-(1:2)], "-", names(wide)[2]),
    W          = sapply(diffs, function(x) unname(shapiro.test(x)$statistic)),
    p          = sapply(diffs, function(x) shapiro.test(x)$p.value),
    row.names  = NULL
  )

  list(
    aov      = anova_test(d, dv = y, wid = ID, within = grp),
    friedman = friedman_test(d, y ~ grp | ID),
    kendall  = friedman_effsize(d, y ~ grp | ID),
    t_ph     = pairwise_t_test(d, y ~ grp, paired = TRUE, p.adjust.method = "holm"),
    w_ph     = pairwise_wilcox_test(d, y ~ grp, paired = TRUE, p.adjust.method = "holm"),
    aov_log  = anova_test(mutate(d, y = log(y)), dv = y, wid = ID, within = grp),
    shapiro  = shapiro
  )
}

# Descriptives in the same shape as the tSNR summary tables.
fd_summary <- function(d) {
  d |>
    group_by(grp) |>
    summarise(
      n      = dplyr::n(),
      mean   = mean(y), sd = sd(y),
      median = median(y), IQR = IQR(y),
      min    = min(y), max = max(y),
      .groups = "drop"
    ) |>
    mutate(across(mean:max, ~ round(.x, 4)))
}

# Percent change relative to the first level.
fd_pct <- function(d) {
  d |>
    group_by(grp) |>
    summarise(mean = mean(y), .groups = "drop") |>
    mutate(`change vs first (%)` = round(100 * (mean - first(mean)) / first(mean), 2),
           mean = round(mean, 4))
}

# Caption noting how many comparisons were annotated.
brk_caption <- function(n_sub, pw, extra = "") {
  n_sig <- sum(pw$p.adj < 0.05)
  paste0(
    "n = ", n_sub, " subjects with complete data.", extra, "\n",
    "Brackets: Holm-corrected paired t-tests; unmarked = ns.\n",
    if (n_sig == 0) "No comparison was significant, so no bracket is drawn."
    else "*** p < .001    ** p < .01    * p < .05"
  )
}

day_runs   <- c(1, 2, 6)
day_labels <- c("Day 1", "Day 2", "Day 3")
d2_runs    <- 2:5
d2_labels  <- paste("Rest", d2_runs)
```

# Part 1: across days (runs 1, 2, 6)

```{r p1-prep}
mean_day <- prep_fd(fd, "mean_FD", day_runs, day_labels)
max_day  <- prep_fd(fd, "max_FD",  day_runs, day_labels)

cat("Complete cases: ", n_distinct(mean_day$ID), " of ",
    attr(mean_day, "n_all"), " subjects\n", sep = "")
```

Subjects excluded, and which of the three runs they have:

```{r p1-dropped}
fd |>
  filter(run %in% day_runs) |>
  group_by(ID) |>
  filter(dplyr::n() < length(day_runs)) |>
  summarise(runs_present = paste(sort(unique(run)), collapse = ", "), .groups = "drop")
```

## Mean FD

```{r p1-mean-summary}
fd_summary(mean_day)
```

```{r p1-mean-fit}
f_mean_day <- fd_fit(mean_day)
```

Repeated-measures ANOVA, factor **day** (3 levels), Greenhouse-Geisser applied
automatically where Mauchly's test indicates a sphericity violation:

```{r p1-mean-anova}
get_anova_table(f_mean_day$aov, correction = "auto")

f_mean_day$aov$`Mauchly's Test for Sphericity`
```

Normality of the paired differences, and the rank-based equivalent:

```{r p1-mean-np}
f_mean_day$shapiro

f_mean_day$friedman

f_mean_day$kendall
```

Post-hoc comparisons, Holm-corrected — parametric and rank-based side by side:

```{r p1-mean-posthoc}
f_mean_day$t_ph |> select(group1, group2, statistic, df, p, p.adj, p.adj.signif)

f_mean_day$w_ph |> select(group1, group2, statistic, p, p.adj, p.adj.signif)
```

```{r p1-mean-plot, fig.width=8, fig.height=5.5}
p_mean_day <- qc_panel(
  mean_day,
  y_lab    = "Mean FD (mm)",
  x_title  = "Acquisition day",
  title    = "Mean framewise displacement by acquisition day",
  subtitle = "Each dot is one subject; Day 2 is the session acquired with EEG",
  caption  = brk_caption(n_distinct(mean_day$ID), f_mean_day$t_ph),
  pw       = f_mean_day$t_ph
)

p_mean_day
```

```{r p1-mean-save, include=FALSE}
save_fig("fd_mean_boxplot", p_mean_day)
```

```{r p1-mean-pct}
fd_pct(mean_day)
```

Log check, since a proportional change is the more natural model for a positive
quantity:

```{r p1-mean-log}
get_anova_table(f_mean_day$aov_log, correction = "auto")
```

## Max FD

```{r p1-max-summary}
fd_summary(max_day)
```

```{r p1-max-fit}
f_max_day <- fd_fit(max_day)
```

```{r p1-max-anova}
get_anova_table(f_max_day$aov, correction = "auto")

f_max_day$aov$`Mauchly's Test for Sphericity`
```

```{r p1-max-np}
f_max_day$shapiro

f_max_day$friedman

f_max_day$kendall
```

```{r p1-max-posthoc}
f_max_day$t_ph |> select(group1, group2, statistic, df, p, p.adj, p.adj.signif)

f_max_day$w_ph |> select(group1, group2, statistic, p, p.adj, p.adj.signif)
```

The y axis is logarithmic here: max FD spans more than an order of magnitude
across subjects (`r round(min(max_day$y), 2)` to `r round(max(max_day$y), 2)` mm),
and on a linear axis a single 5 mm outlier compresses every box into the bottom
fifth of the panel. The boxes and points are the same data as everywhere else in
this report, only the scale differs.

```{r p1-max-plot, fig.width=8, fig.height=5.5}
p_max_day <- qc_panel(
  max_day,
  y_lab    = "Max FD (mm, log scale)",
  x_title  = "Acquisition day",
  title    = "Maximal framewise displacement by acquisition day",
  subtitle = "Each dot is one subject; Day 2 is the session acquired with EEG",
  caption  = brk_caption(n_distinct(max_day$ID), f_max_day$t_ph),
  pw       = f_max_day$t_ph,
  log_y    = TRUE
)

p_max_day
```

```{r p1-max-save, include=FALSE}
save_fig("fd_max_boxplot", p_max_day)
```

```{r p1-max-pct}
fd_pct(max_day)
```

```{r p1-max-log}
get_anova_table(f_max_day$aov_log, correction = "auto")
```

# Part 2: within Day 2 (runs 2-5)

```{r p2-prep}
mean_d2 <- prep_fd(fd, "mean_FD", d2_runs, d2_labels)
max_d2  <- prep_fd(fd, "max_FD",  d2_runs, d2_labels)

cat("Complete cases: ", n_distinct(mean_d2$ID), " of ",
    attr(mean_d2, "n_all"), " subjects with any Day 2 run\n", sep = "")
```

```{r p2-dropped}
fd |>
  filter(run %in% d2_runs) |>
  group_by(ID) |>
  filter(dplyr::n() < length(d2_runs)) |>
  summarise(runs_present = paste(sort(unique(run)), collapse = ", "), .groups = "drop")
```

## Mean FD

```{r p2-mean-summary}
fd_summary(mean_d2)
```

```{r p2-mean-fit}
f_mean_d2 <- fd_fit(mean_d2)
```

```{r p2-mean-anova}
get_anova_table(f_mean_d2$aov, correction = "auto")

f_mean_d2$aov$`Mauchly's Test for Sphericity`
```

```{r p2-mean-np}
f_mean_d2$shapiro

f_mean_d2$friedman

f_mean_d2$kendall
```

```{r p2-mean-posthoc}
f_mean_d2$t_ph |> select(group1, group2, statistic, df, p, p.adj, p.adj.signif)

f_mean_d2$w_ph |> select(group1, group2, statistic, p, p.adj, p.adj.signif)
```

```{r p2-mean-plot, fig.width=8, fig.height=5.5}
p_mean_d2 <- qc_panel(
  mean_d2,
  y_lab    = "Mean FD (mm)",
  x_title  = "Resting-state run (Day 2)",
  title    = "Mean FD across the four EEG runs of Day 2",
  subtitle = "Each dot is one subject; all four runs acquired with the EEG cap in place",
  caption  = brk_caption(n_distinct(mean_d2$ID), f_mean_d2$t_ph),
  pw       = f_mean_d2$t_ph,
  short    = \(s) sub("^Rest ", "R", s)
)

p_mean_d2
```

```{r p2-mean-save, include=FALSE}
save_fig("fd_mean_boxplot_day2", p_mean_d2)
```

```{r p2-mean-pct}
fd_pct(mean_d2)
```

```{r p2-mean-log}
get_anova_table(f_mean_d2$aov_log, correction = "auto")
```

## Max FD

```{r p2-max-summary}
fd_summary(max_d2)
```

```{r p2-max-fit}
f_max_d2 <- fd_fit(max_d2)
```

```{r p2-max-anova}
get_anova_table(f_max_d2$aov, correction = "auto")

f_max_d2$aov$`Mauchly's Test for Sphericity`
```

```{r p2-max-np}
f_max_d2$shapiro

f_max_d2$friedman

f_max_d2$kendall
```

```{r p2-max-posthoc}
f_max_d2$t_ph |> select(group1, group2, statistic, df, p, p.adj, p.adj.signif)

f_max_d2$w_ph |> select(group1, group2, statistic, p, p.adj, p.adj.signif)
```

```{r p2-max-plot, fig.width=8, fig.height=5.5}
p_max_d2 <- qc_panel(
  max_d2,
  y_lab    = "Max FD (mm, log scale)",
  x_title  = "Resting-state run (Day 2)",
  title    = "Max FD across the four EEG runs of Day 2",
  subtitle = "Each dot is one subject; all four runs acquired with the EEG cap in place",
  caption  = brk_caption(n_distinct(max_d2$ID), f_max_d2$t_ph),
  pw       = f_max_d2$t_ph,
  log_y    = TRUE,
  short    = \(s) sub("^Rest ", "R", s)
)

p_max_d2
```

```{r p2-max-save, include=FALSE}
save_fig("fd_max_boxplot_day2", p_max_d2)
```

```{r p2-max-pct}
fd_pct(max_d2)
```

```{r p2-max-log}
get_anova_table(f_max_d2$aov_log, correction = "auto")
```

# Summary of observed effects

```{r summary-numbers, include=FALSE}
gv  <- function(fit, col, eff = "grp") {
  tab <- get_anova_table(fit, correction = "auto")
  tab[[col]][tab$Effect == eff]
}
fmt_p <- function(p) ifelse(p < 0.001, "< .001",
                            paste0("= ", format(round(p, 3), nsmall = 3)))
gm <- function(d, g) round(mean(d$y[d$grp == g]), 3)

n_day <- n_distinct(mean_day$ID)
n_d2  <- n_distinct(mean_d2$ID)

# How many of all the post-hoc comparisons in the report survive correction.
all_ph <- bind_rows(
  f_mean_day$t_ph, f_max_day$t_ph, f_mean_d2$t_ph, f_max_d2$t_ph,
  f_mean_day$w_ph, f_max_day$w_ph, f_mean_d2$w_ph, f_max_d2$w_ph
)
n_ph_sig <- sum(all_ph$p.adj < 0.05)

min_shapiro_max <- min(c(f_max_day$shapiro$p, f_max_d2$shapiro$p))
```

**Head motion does not differ between the three acquisition days.** Mean FD is
`r gm(mean_day, "Day 1")` mm (Day 1), `r gm(mean_day, "Day 2")` mm (Day 2) and
`r gm(mean_day, "Day 3")` mm (Day 3). The ANOVA is nominally significant
(*F*(`r round(gv(f_mean_day$aov,"DFn"),2)`, `r round(gv(f_mean_day$aov,"DFd"),2)`) =
`r round(gv(f_mean_day$aov,"F"),2)`, *p* `r fmt_p(gv(f_mean_day$aov,"p"))`,
ges = `r round(gv(f_mean_day$aov,"ges"),3)`), but that result does not hold up:
no post-hoc comparison survives Holm correction (smallest
*p*<sub>adj</sub> = `r round(min(f_mean_day$t_ph$p.adj), 3)`), Friedman's test on
the same data gives *p* = `r round(f_mean_day$friedman$p, 3)` with Kendall's
*W* = `r round(f_mean_day$kendall$effsize, 3)` (small), and the effect drops to
*p* `r fmt_p(gv(f_mean_day$aov_log,"p"))` on the log scale. The nominal effect is
carried by Day 3 being about
`r round(fd_pct(mean_day)$"change vs first (%)"[3], 0)`% higher than Day 1, not by
the EEG day. Max FD shows nothing at all
(*F*(`r round(gv(f_max_day$aov,"DFn"),2)`, `r round(gv(f_max_day$aov,"DFd"),2)`) =
`r round(gv(f_max_day$aov,"F"),2)`, *p* `r fmt_p(gv(f_max_day$aov,"p"))`;
Friedman *p* = `r round(f_max_day$friedman$p, 3)`).

**Motion is also flat across the four EEG runs of Day 2.** Mean FD declines
slightly and monotonically after the first run — `r gm(mean_d2, "Rest 2")`,
`r gm(mean_d2, "Rest 3")`, `r gm(mean_d2, "Rest 4")`, `r gm(mean_d2, "Rest 5")` mm
— but neither the ANOVA
(*F*(`r round(gv(f_mean_d2$aov,"DFn"),2)`, `r round(gv(f_mean_d2$aov,"DFd"),2)`) =
`r round(gv(f_mean_d2$aov,"F"),2)`, *p* `r fmt_p(gv(f_mean_d2$aov,"p"))`,
ges = `r round(gv(f_mean_d2$aov,"ges"),3)`) nor Friedman's test
(*p* = `r round(f_mean_d2$friedman$p, 3)`) reaches significance, and no post-hoc
survives correction. Max FD is equally null
(*p* `r fmt_p(gv(f_max_d2$aov,"p"))`; Friedman *p* =
`r round(f_max_d2$friedman$p, 3)`). There is no evidence of motion drift over the
EEG session, and no evidence that subjects settle or deteriorate between runs.

**In total, `r n_ph_sig` of `r nrow(all_ph)` post-hoc comparisons in this report
survive Holm correction**, which is why none of the four figures carries a
significance bracket.

**A caveat on max FD.** The paired differences in max FD are strongly
non-normal — Shapiro-Wilk *p* as low as
`r format(signif(min_shapiro_max, 2), scientific = TRUE)` — because the measure
is the largest of ~330 values and is dominated by isolated events (one subject
reaches `r round(max(max_day$y), 1)` mm on Day 3, five times the group median).
For max FD the Friedman and Wilcoxon results are therefore the primary ones, not
the ANOVA. They agree with it here, so the conclusion is unaffected, but a future
analysis with a real effect should not lean on the parametric test for this
variable. Mean FD is much better behaved (Shapiro-Wilk *p* >
`r round(min(f_mean_day$shapiro$p), 2)` for the day differences).

# Does motion explain the Day 2 tSNR loss?

This section is not part of the tSNR analysis being replicated; it is here
because the FD result bears directly on how the tSNR result should be read. The
tSNR report found a ~6% drop in grey-matter tSNR on Day 2, uniform across
echoes. Motion is the obvious alternative explanation — a subject who moves more
with a cap on would show lower tSNR for reasons having nothing to do with the
cap's effect on the signal.

The data above already argues against the simplest version of that: mean FD on
Day 2 (`r gm(mean_day, "Day 2")` mm) is essentially identical to Day 1
(`r gm(mean_day, "Day 1")` mm), *p* =
`r round(f_mean_day$t_ph$p.adj[f_mean_day$t_ph$group1 == "Day 1" & f_mean_day$t_ph$group2 == "Day 2"], 3)`.
A group-level tSNR drop cannot be produced by a group-level motion increase that
did not happen. But that leaves a subtler possibility open: motion might still
explain *which* subjects lost the most tSNR. Testing that needs the two tables
joined.

```{r motion-tsnr}
tsnr_wide <- read.csv("tSNR_results.txt", na.strings = c("NA", "NaN", ""))
names(tsnr_wide)[1] <- "ID"

# tSNR averaged over the four echoes, for runs 1 and 2.
tsnr_delta <- tsnr_wide |>
  transmute(
    ID     = ID,
    tsnr_1 = rowMeans(across(c(r1e1, r1e2, r1e3, r1e4))),
    tsnr_2 = rowMeans(across(c(r2e1, r2e2, r2e3, r2e4)))
  ) |>
  mutate(d_tsnr = tsnr_2 - tsnr_1)

fd_delta <- fd |>
  filter(run %in% c(1, 2)) |>
  select(ID, run, mean_FD) |>
  pivot_wider(names_from = run, values_from = mean_FD, names_prefix = "fd_") |>
  mutate(d_fd = fd_2 - fd_1)

joined <- inner_join(tsnr_delta, fd_delta, by = "ID") |>
  filter(!is.na(d_tsnr), !is.na(d_fd))

cat("Subjects with both measures on runs 1 and 2: ", nrow(joined), "\n", sep = "")
```

The change in motion and the change in tSNR *are* related across subjects:

```{r motion-tsnr-cor}
cor.test(joined$d_fd, joined$d_tsnr, method = "spearman")
```

So motion is not irrelevant. The question is whether it accounts for the *mean*
drop or only for the spread around it. Regressing the tSNR change on the motion
change separates the two: the slope is the motion contribution, and the
intercept is the tSNR change expected for a subject whose motion did not change
at all.

```{r motion-tsnr-lm}
fit_conf <- lm(d_tsnr ~ d_fd, data = joined)
summary(fit_conf)
```

```{r motion-tsnr-plot, fig.width=6.5, fig.height=5}
ggplot(joined, aes(x = d_fd, y = d_tsnr)) +
  geom_hline(yintercept = 0, colour = "grey70", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "grey70", linewidth = 0.4) +
  geom_smooth(method = "lm", formula = y ~ x, colour = "black",
              linewidth = 0.5, fill = "grey85") +
  geom_point(shape = 21, fill = "black", colour = "white", size = 2.6, stroke = 0.4) +
  labs(
    title    = "Change in motion vs change in tSNR, Day 1 to Day 2",
    subtitle = "Each dot is one subject; tSNR averaged over the four echoes",
    x        = "Change in mean FD (mm)",
    y        = "Change in tSNR"
  )
```

```{r motion-tsnr-save, include=FALSE}
save_fig("fd_vs_tsnr_change", last_plot(), h = 13)
```

```{r motion-tsnr-numbers, include=FALSE}
ct     <- cor.test(joined$d_fd, joined$d_tsnr, method = "spearman")
cf     <- summary(fit_conf)$coefficients
b0     <- cf["(Intercept)", "Estimate"];  p0 <- cf["(Intercept)", "Pr(>|t|)"]
b1     <- cf["d_fd", "Estimate"];         p1 <- cf["d_fd", "Pr(>|t|)"]
r2     <- summary(fit_conf)$r.squared
n_up   <- sum(joined$d_fd > 0)
mean_d <- mean(joined$d_tsnr)
```

**Both things are true, and they answer different questions.** Across the
`r nrow(joined)` subjects with both measures, motion change does predict tSNR
change, in the expected direction and with a respectable effect size: Spearman
*rho* = `r round(unname(ct$estimate), 2)` (*p* = `r round(ct$p.value, 3)`), and
in the regression the slope is significant (*b* = `r round(b1, 1)` tSNR units per
mm of FD, *p* = `r round(p1, 3)`, R² = `r round(r2, 2)`). Subjects who moved more
on Day 2 did lose more tSNR. Motion is a genuine contributor to the
*subject-to-subject spread* in the Day 2 loss.

It is not, however, the cause of the loss itself. The regression intercept — the
expected tSNR change for a subject whose motion was identical on the two days —
is `r round(b0, 2)` (*p* = `r format(round(p0, 4), nsmall = 4)`), against a raw
mean change of `r round(mean_d, 2)`. In other words, essentially the entire mean
drop remains at zero motion change. That is the expected result given the
group-level numbers: only `r n_up` of `r nrow(joined)` subjects moved more on Day 2
at all, and mean FD across the group was flat, so motion has no mean shift
available to transmit to tSNR.

The reading in the tSNR report therefore stands, with one refinement. The Day 2
tSNR loss is a property of the EEG hardware being in the bore rather than of
subjects behaving differently while wearing it — but motion still modulates how
large that loss is for a given subject, and a study using per-subject Day 2 tSNR
as a covariate would do well to include mean FD alongside it. Two caveats on this
section specifically: n = `r nrow(joined)` is small for a correlation, so the
slope estimate is imprecise; and FD and tSNR are not independent measurements,
since motion enters the tSNR denominator directly through the temporal standard
deviation.

## Session info

```{r session}
sessionInfo()
```
