---
title: "tSNR across days and echoes"
author: "Benjamin Meyer"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 8, fig.height = 5.5, dpi = 200,
  dev.args = list(bg = "white")
)
library(ggplot2)
library(tidyr)
library(dplyr)
library(rstatix)
library(knitr)
```

## Read the data

The file is comma-separated: column 1 is the subject ID, the remaining columns
are tSNR values named `r<run>e<echo>` (R = resting-state scan, E = echo).
`NaN` entries (runs/echoes missing for a subject) are read as `NA`.

```{r read}
tsnr <- read.csv(
  "tSNR_results.txt",
  header     = TRUE,
  na.strings = c("NA", "NaN", "")
)

names(tsnr)[1] <- "ID"
value_cols <- setdiff(names(tsnr), "ID")

str(tsnr)
```

## Reshape to long format

`run` and `echo` are pulled out of the column names.

```{r reshape}
tsnr_long <- tsnr |>
  pivot_longer(
    cols      = all_of(value_cols),
    names_to  = "scan",
    values_to = "tSNR"
  ) |>
  mutate(
    scan = factor(scan, levels = value_cols),
    run  = as.integer(sub("^r(\\d+)e\\d+$", "\\1", scan)),
    echo = factor(as.integer(sub("^r\\d+e(\\d+)$", "\\1", scan)))
  ) |>
  filter(!is.na(tSNR))

head(tsnr_long)
```

## Run-to-day mapping

Taken from the acquisition loop in the data-generating scripts
(`eeg_fmri_ext_SNR_1_v2.m:63` and `eeg_fmri_ext_SNR_2_v2.sh:61`, both
`rest_day = [1 2 2 2 2 3]`), with the EEG status from
`eeg_fmri_ext_SNR_1_v2.m:62` ("1, 6: without EEG; 2, 3, 4, 5: with EEG"):

| Run | Session | EEG |
|-----|---------|-----|
| 1 | `ses-Day1_task-rest1` | no |
| 2 | `ses-Day2_task-rest2` | **yes** |
| 3 | `ses-Day2_task-rest3` | **yes** |
| 4 | `ses-Day2_task-rest4` | **yes** |
| 5 | `ses-Day2_task-rest5` | **yes** |
| 6 | `ses-Day3_task-rest6` | no |

Two questions follow from this, and they are analysed separately below:

1. **Across days** — runs 1, 2 and 6, i.e. the first resting-state scan of each
   day, one run per day. This is the session-stability question.
2. **Within Day 2** — runs 2 to 5, the four EEG runs of the middle session.
   This asks whether tSNR drifts over the course of the EEG session.

## Helpers: plotting and significance annotation

Both analyses produce the same kind of figure (echo blocks on the x-axis, the
grouping factor within each block) and the same kind of post-hoc table, so the
plot and the significance brackets are written once here and reused.

Brackets are drawn only for the **within-echo** comparisons that survive Holm
correction: `***` *p* < .001, `**` *p* < .01, `*` *p* < .05. Non-significant
comparisons are left unmarked. Each bracket sits just above the data of its own
echo block rather than at a single panel-wide height — echo 1 has roughly three
times the tSNR of echo 4, so one common height would strand the echo-4 brackets
far above their boxes.

```{r helpers}
p_stars <- function(p) {
  ifelse(p < 0.001, "***",
  ifelse(p < 0.01,  "**",
  ifelse(p < 0.05,  "*", NA_character_)))
}

# Lowest free vertical tier per bracket, so that overlapping spans stack
# instead of being drawn on top of each other.
pack_tiers <- function(xmin, xmax, pad = 0.35) {
  tier <- integer(length(xmin))
  for (i in seq_along(xmin)) {
    j <- seq_len(i - 1L)
    t <- 1L
    while (any(tier[j] == t & xmin[i] <= xmax[j] + pad & xmax[i] >= xmin[j] - pad)) {
      t <- t + 1L
    }
    tier[i] <- t
  }
  tier
}

# x-axis cells: one box per echo x group combination, ordered echo-major so the
# groups being compared sit next to each other.
add_cells <- function(d) {
  d   <- d |> mutate(echo = droplevels(echo), grp = droplevels(grp))
  lev <- paste0(rep(levels(d$echo), each = nlevels(d$grp)), "_", levels(d$grp))
  d |> mutate(
    cell = factor(paste0(as.character(echo), "_", as.character(grp)), levels = lev)
  )
}

# pw: within-echo pairwise table with columns echo, group1, group2, p.adj.
# Returns a list of layers, empty if nothing is significant.
sig_geom <- function(pw, d, step_frac = 0.055, tick_frac = 0.014) {
  grp  <- levels(d$grp)
  span <- diff(range(d$tSNR))
  step <- step_frac * span
  tick <- tick_frac * span

  y_echo <- d |>
    group_by(echo) |>
    summarise(y0 = max(tSNR), .groups = "drop") |>
    mutate(echo = as.character(echo))

  b <- pw |>
    ungroup() |>
    mutate(label = p_stars(p.adj), echo = as.character(echo)) |>
    filter(!is.na(label)) |>
    mutate(
      e    = match(echo, levels(d$echo)),
      xmin = (e - 1) * length(grp) + pmin(match(group1, grp), match(group2, grp)),
      xmax = (e - 1) * length(grp) + pmax(match(group1, grp), match(group2, grp))
    ) |>
    left_join(y_echo, by = "echo") |>
    arrange(e, xmax - xmin, xmin) |>
    group_by(e) |>
    mutate(tier = pack_tiers(xmin, xmax)) |>
    ungroup() |>
    mutate(y = y0 + tier * step)

  if (nrow(b) == 0) return(list())

  # Four points per bracket: down-tick, across, down-tick.
  paths <- data.frame(
    id = rep(seq_len(nrow(b)), each = 4),
    x  = as.vector(rbind(b$xmin, b$xmin, b$xmax, b$xmax)),
    y  = as.vector(rbind(b$y - tick, b$y, b$y, b$y - tick))
  )

  list(
    geom_path(
      data = paths, aes(x = x, y = y, group = id),
      colour = "black", linewidth = 0.4, inherit.aes = FALSE
    ),
    geom_text(
      data = b, aes(x = (xmin + xmax) / 2, y = y, label = label),
      vjust = -0.15, size = 4.6, colour = "black", inherit.aes = FALSE
    ),
    expand_limits(y = max(b$y) + 1.8 * step)
  )
}
```

Black boxes and points, no background grid, explicit x- and y-axis lines, and
`base_size = 14` so the figure stays legible at A4 width. The x-axis is
two-tiered: the grouping factor on the axis itself, echo beneath.

```{r theme}
base_font <- 14   # readable when the 8 x 5.5 in figure is placed on A4

theme_tsnr <- theme_classic(base_size = base_font) +
  theme(
    panel.grid       = element_blank(),
    axis.line        = element_line(colour = "black", linewidth = 0.5),
    axis.ticks       = element_line(colour = "black", linewidth = 0.5),
    axis.text        = element_text(colour = "black"),
    axis.title       = element_text(colour = "black"),
    axis.title.x     = element_text(margin = margin(t = 30)),
    axis.title.y     = element_text(margin = margin(r = 8)),
    plot.title       = element_text(face = "bold", size = base_font + 1),
    plot.subtitle    = element_text(size = base_font - 2),
    plot.caption     = element_text(size = base_font - 3, hjust = 0, colour = "grey30"),
    plot.margin      = margin(6, 8, 14, 6)
  )

theme_set(theme_tsnr)
```

```{r panel-fn}
tsnr_panel <- function(d, pw = NULL, title, subtitle, caption, x_title,
                       short = identity) {
  n_g <- nlevels(d$grp)

  # One label per echo, centred under that echo's block of boxes.
  echo_tier <- data.frame(
    x     = (seq_len(nlevels(d$echo)) - 1) * n_g + (n_g + 1) / 2,
    label = paste0("Echo ", levels(d$echo))
  )

  # Jitter is horizontal only, so no point is drawn at a false tSNR value.
  p <- ggplot(d, aes(x = cell, y = tSNR)) +
    geom_boxplot(
      fill = "white", colour = "black",
      linewidth = 0.5, width = 0.62,
      outlier.shape = NA          # outliers are already shown as points below
    ) +
    geom_point(
      position = position_jitter(width = 0.16, height = 0, seed = 42),
      shape = 21, fill = "black", colour = "white",
      size = 2.1, stroke = 0.35, alpha = 0.9
    ) +
    geom_text(
      data = echo_tier, aes(x = x, y = -Inf, label = label),
      vjust = 3.5, size = base_font / .pt, colour = "black",
      inherit.aes = FALSE
    ) +
    scale_x_discrete(labels = \(z) short(sub("^[^_]+_", "", z))) +
    scale_y_continuous(expand = expansion(mult = c(0.04, 0.06))) +
    coord_cartesian(clip = "off") +
    labs(
      title = title, subtitle = subtitle, caption = caption,
      x = x_title, y = "tSNR (grey matter)"
    )

  if (!is.null(pw)) p <- p + sig_geom(pw, d)
  p
}
```

# Part 1: across days (runs 1, 2, 6)

```{r subset}
runs_keep <- c(1, 2, 6)
day_labels <- c("Day 1", "Day 2", "Day 3")

tsnr_day <- tsnr_long |>
  filter(run %in% runs_keep) |>
  mutate(
    day  = factor(run, levels = runs_keep, labels = day_labels),
    scan = droplevels(scan),
    echo = droplevels(echo)
  )

# The ANOVA is a fully within-subject design, so it needs subjects with a value
# in every day x echo cell.
n_cells <- length(runs_keep) * nlevels(tsnr_day$echo)

tsnr_rm <- tsnr_day |>
  group_by(ID) |>
  filter(dplyr::n() == n_cells) |>
  ungroup() |>
  mutate(ID = factor(ID), grp = day) |>
  add_cells()

cat(
  "Complete cases: ", n_distinct(tsnr_rm$ID), " of ",
  n_distinct(tsnr_day$ID), " subjects\n", sep = ""
)
```

Subjects excluded, and which of the three runs they have:

```{r dropped}
tsnr_day |>
  group_by(ID) |>
  filter(dplyr::n() < n_cells) |>
  summarise(runs_present = paste(sort(unique(run)), collapse = ", "), .groups = "drop")
```

## Summary statistics

```{r summary-table}
tsnr_rm |>
  group_by(echo, day) |>
  summarise(
    n      = dplyr::n(),
    mean   = mean(tSNR),
    sd     = sd(tSNR),
    median = median(tSNR),
    IQR    = IQR(tSNR),
    min    = min(tSNR),
    max    = max(tSNR),
    .groups = "drop"
  ) |>
  mutate(across(mean:max, ~ round(.x, 2)))
```

## Two-way repeated-measures ANOVA

Every subject contributes a value in all `r n_cells` cells, so **day** (3 levels)
and **echo** (4 levels) are both within-subject factors. One ANOVA, with both
main effects and their interaction. Greenhouse-Geisser is applied automatically
where Mauchly's test indicates a sphericity violation; `ges` is generalised
eta-squared.

```{r anova}
fit <- anova_test(tsnr_rm, dv = tSNR, wid = ID, within = c(day, echo))

get_anova_table(fit, correction = "auto")
```

Sphericity:

```{r sphericity}
fit$`Mauchly's Test for Sphericity`

fit$`Sphericity Corrections`
```

### Marginal means

```{r marginals}
tsnr_rm |>
  group_by(day) |>
  summarise(mean = round(mean(tSNR), 2), sd = round(sd(tSNR), 2), .groups = "drop")

tsnr_rm |>
  group_by(echo) |>
  summarise(mean = round(mean(tSNR), 2), sd = round(sd(tSNR), 2), .groups = "drop")
```

### Post-hoc comparisons

Paired *t*-tests between days, Holm-corrected. First as a follow-up to the day
main effect, then within each echo as a follow-up to the interaction.

For the main-effect follow-up the four echoes of a subject are averaged into one
value per subject and day *before* the test. Testing the un-averaged rows would
treat each echo as an independent replicate and give *df* =
`r nlevels(tsnr_rm$echo) * n_distinct(tsnr_rm$ID) - 1`
instead of `r n_distinct(tsnr_rm$ID) - 1`, which inflates the *t* statistics.

```{r posthoc-day}
day_posthoc <- tsnr_rm |>
  group_by(ID, day) |>
  summarise(tSNR = mean(tSNR), .groups = "drop") |>
  pairwise_t_test(tSNR ~ day, paired = TRUE, p.adjust.method = "holm")

day_posthoc |>
  select(group1, group2, statistic, df, p, p.adj, p.adj.signif)
```

Within each echo — these are the comparisons annotated on the figure:

```{r posthoc-echo}
day_by_echo <- tsnr_rm |>
  group_by(echo) |>
  pairwise_t_test(tSNR ~ day, paired = TRUE, p.adjust.method = "holm")

day_by_echo |>
  select(echo, group1, group2, statistic, df, p, p.adj, p.adj.signif)
```

## Plot

```{r plot, fig.width=8, fig.height=5.5}
p_tsnr <- tsnr_panel(
  tsnr_rm,
  pw       = day_by_echo,
  title    = "Temporal SNR by echo and acquisition day",
  subtitle = "Each dot is one subject; Day 2 is the session acquired with EEG",
  x_title  = "Acquisition day",
  caption  = paste0(
    "n = ", n_distinct(tsnr_rm$ID), " subjects with complete data.\n",
    "Brackets: Holm-corrected paired t-tests within echo; unmarked = ns.\n",
    "*** p < .001    ** p < .01    * p < .05"
  ),
  short = \(s) sub("^Day ", "D", s)
)

p_tsnr
```

```{r plot-save, include=FALSE}
# Vector copy at A4 text width, for inclusion in a manuscript.
ggsave("tSNR_boxplot.pdf", p_tsnr, width = 17, height = 12, units = "cm")
```

### Size of the day differences, per echo

The interaction is easier to read as a proportion, because tSNR itself differs
by a factor of three across echoes.

```{r pct-change}
pct_day <- tsnr_rm |>
  group_by(echo, day) |>
  summarise(mean = mean(tSNR), .groups = "drop") |>
  pivot_wider(names_from = day, values_from = mean) |>
  mutate(
    `Day 2 vs Day 1 (%)` = 100 * (`Day 2` - `Day 1`) / `Day 1`,
    `Day 3 vs Day 1 (%)` = 100 * (`Day 3` - `Day 1`) / `Day 1`
  )

pct_day |>
  mutate(across(where(is.numeric), ~ round(.x, 2)))
```

### Is the interaction just a scaling effect?

An equal *proportional* change across echoes would still show up as an
interaction in absolute units, since echo 1 has three times the tSNR of echo 4.
Refitting on `log(tSNR)` turns proportional effects into additive ones, so if
the interaction is purely a scaling artefact it disappears here.

```{r log-check}
fit_log <- tsnr_rm |>
  mutate(log_tSNR = log(tSNR)) |>
  anova_test(dv = log_tSNR, wid = ID, within = c(day, echo))

get_anova_table(fit_log, correction = "auto")
```

# Part 2: within Day 2 (runs 2-5)

All four runs of the Day 2 session were acquired with the EEG cap in place. If
the cap, or cap-related drift (gel drying, subject settling, heating), degraded
tSNR progressively, it should show up as a run effect here.

```{r day2-subset}
runs_d2   <- 2:5
run_labels <- paste("Rest", runs_d2)

tsnr_d2_all <- tsnr_long |>
  filter(run %in% runs_d2) |>
  mutate(
    run_f = factor(run, levels = runs_d2, labels = run_labels),
    scan  = droplevels(scan),
    echo  = droplevels(echo)
  )

n_cells_d2 <- length(runs_d2) * nlevels(tsnr_d2_all$echo)

tsnr_d2 <- tsnr_d2_all |>
  group_by(ID) |>
  filter(dplyr::n() == n_cells_d2) |>
  ungroup() |>
  mutate(ID = factor(ID), grp = run_f) |>
  add_cells()

cat(
  "Complete cases: ", n_distinct(tsnr_d2$ID), " of ",
  n_distinct(tsnr_d2_all$ID), " subjects with any Day 2 run\n", sep = ""
)
```

```{r day2-dropped}
tsnr_d2_all |>
  group_by(ID) |>
  filter(dplyr::n() < n_cells_d2) |>
  summarise(runs_present = paste(sort(unique(run)), collapse = ", "), .groups = "drop")
```

## Summary statistics

```{r day2-summary}
tsnr_d2 |>
  group_by(echo, run_f) |>
  summarise(
    n      = dplyr::n(),
    mean   = mean(tSNR),
    sd     = sd(tSNR),
    median = median(tSNR),
    IQR    = IQR(tSNR),
    min    = min(tSNR),
    max    = max(tSNR),
    .groups = "drop"
  ) |>
  mutate(across(mean:max, ~ round(.x, 2)))
```

## Two-way repeated-measures ANOVA

**run** (4 levels: rest2-rest5) and **echo** (4 levels) are both within-subject.
Same model as Part 1, with run in place of day.

```{r day2-anova}
fit_d2 <- anova_test(tsnr_d2, dv = tSNR, wid = ID, within = c(run_f, echo))

get_anova_table(fit_d2, correction = "auto")
```

Sphericity:

```{r day2-sphericity}
fit_d2$`Mauchly's Test for Sphericity`

fit_d2$`Sphericity Corrections`
```

### Marginal means

```{r day2-marginals}
tsnr_d2 |>
  group_by(run_f) |>
  summarise(mean = round(mean(tSNR), 2), sd = round(sd(tSNR), 2), .groups = "drop")

tsnr_d2 |>
  group_by(echo) |>
  summarise(mean = round(mean(tSNR), 2), sd = round(sd(tSNR), 2), .groups = "drop")
```

### Post-hoc comparisons

Reported for completeness even though the omnibus run effect is not significant,
so they should be read as descriptive rather than confirmatory. Subject means
across echoes for the main-effect follow-up, as in Part 1.

```{r day2-posthoc-run}
tsnr_d2 |>
  group_by(ID, run_f) |>
  summarise(tSNR = mean(tSNR), .groups = "drop") |>
  pairwise_t_test(tSNR ~ run_f, paired = TRUE, p.adjust.method = "holm") |>
  select(group1, group2, statistic, df, p, p.adj, p.adj.signif)
```

Within each echo — these are the comparisons annotated on the Day 2 figure:

```{r day2-posthoc-echo}
run_by_echo <- tsnr_d2 |>
  group_by(echo) |>
  pairwise_t_test(tSNR ~ run_f, paired = TRUE, p.adjust.method = "holm")

run_by_echo |>
  select(echo, group1, group2, statistic, df, p, p.adj, p.adj.signif)
```

```{r day2-nsig, echo=FALSE}
n_sig_d2 <- sum(run_by_echo$p.adj < 0.05)
cat("Within-echo run comparisons surviving Holm correction: ", n_sig_d2,
    " of ", nrow(run_by_echo), "\n", sep = "")
```

## Plot

```{r day2-plot, fig.width=8, fig.height=5.5}
p_d2 <- tsnr_panel(
  tsnr_d2,
  pw       = run_by_echo,
  title    = "Temporal SNR across the four EEG runs of Day 2",
  subtitle = "Each dot is one subject; all four runs acquired with the EEG cap in place",
  x_title  = "Resting-state run (Day 2)",
  caption  = paste0(
    "n = ", n_distinct(tsnr_d2$ID), " subjects with all four Day 2 runs.\n",
    "Brackets: Holm-corrected paired t-tests within echo; unmarked = ns.\n",
    if (n_sig_d2 == 0) "No within-echo comparison was significant, so no bracket is drawn."
    else "*** p < .001    ** p < .01    * p < .05"
  ),
  short = \(s) sub("^Rest ", "R", s)
)

p_d2
```

```{r day2-plot-save, include=FALSE}
ggsave("tSNR_boxplot_day2.pdf", p_d2, width = 17, height = 12, units = "cm")
```

### Size of the run differences, per echo

Relative to rest2, the first EEG run of the session.

```{r day2-pct}
pct_d2 <- tsnr_d2 |>
  group_by(echo, run_f) |>
  summarise(mean = mean(tSNR), .groups = "drop") |>
  pivot_wider(names_from = run_f, values_from = mean) |>
  mutate(
    `Rest 3 vs Rest 2 (%)` = 100 * (`Rest 3` - `Rest 2`) / `Rest 2`,
    `Rest 4 vs Rest 2 (%)` = 100 * (`Rest 4` - `Rest 2`) / `Rest 2`,
    `Rest 5 vs Rest 2 (%)` = 100 * (`Rest 5` - `Rest 2`) / `Rest 2`
  )

pct_d2 |>
  mutate(across(where(is.numeric), ~ round(.x, 2)))
```

### Log check

```{r day2-log-check}
tsnr_d2 |>
  mutate(log_tSNR = log(tSNR)) |>
  anova_test(dv = log_tSNR, wid = ID, within = c(run_f, echo)) |>
  get_anova_table(correction = "auto")
```

# Summary of observed effects

```{r summary-numbers, include=FALSE}
at    <- get_anova_table(fit, correction = "auto")
atl   <- get_anova_table(fit_log, correction = "auto")
at2   <- get_anova_table(fit_d2, correction = "auto")
gv    <- function(tab, eff, col) tab[[col]][tab$Effect == eff]
n_sub <- n_distinct(tsnr_rm$ID)

em    <- function(e) round(mean(tsnr_rm$tSNR[tsnr_rm$echo == e]), 1)
dm    <- function(d) round(mean(tsnr_rm$tSNR[tsnr_rm$day  == d]), 1)
rm2m  <- function(r) round(mean(tsnr_d2$tSNR[tsnr_d2$run_f == r]), 1)

# Day 2 cost, as a percentage of the mean of the two non-EEG days.
d2_cost <- 100 * (dm("Day 2") - mean(c(dm("Day 1"), dm("Day 3")))) /
  mean(c(dm("Day 1"), dm("Day 3")))
pct_rng <- range(pct_day$`Day 2 vs Day 1 (%)`)
d13_rng <- range(abs(pct_day$`Day 3 vs Day 1 (%)`))
fmt_p   <- function(p) ifelse(p < 0.001, "< .001", paste0("= ", format(round(p, 3), nsmall = 3)))
```

**Echo dominates.** tSNR falls monotonically with echo number — `r em(1)`
(echo 1), `r em(2)`, `r em(3)`, and `r em(4)` (echo 4) — a very large effect
(*F*(`r round(gv(at,"echo","DFn"),2)`, `r round(gv(at,"echo","DFd"),2)`) =
`r round(gv(at,"echo","F"),1)`, *p* < .001,
ges = `r round(gv(at,"echo","ges"),3)`). This is the expected T2\* decay: later
echoes have longer TE, so less signal remains while thermal noise stays roughly
constant. It is a property of the sequence, not a data-quality finding.

**Day 2 loses about `r round(abs(d2_cost), 0)`% of its tSNR.** Collapsed across
echoes, mean tSNR is `r dm("Day 1")` (Day 1), `r dm("Day 2")` (Day 2) and
`r dm("Day 3")` (Day 3):
*F*(`r round(gv(at,"day","DFn"),2)`, `r round(gv(at,"day","DFd"),2)`) =
`r round(gv(at,"day","F"),2)`, *p* `r fmt_p(gv(at,"day","p"))`,
ges = `r round(gv(at,"day","ges"),3)`. The post-hoc pattern is unambiguous:
Day 1 > Day 2 and Day 3 > Day 2 both survive correction, while Day 1 vs Day 3 is
nowhere near significance (*p* =
`r format(round(day_posthoc$p.adj[day_posthoc$group1 == "Day 1" & day_posthoc$group2 == "Day 3"], 2), nsmall = 2)`).
Day 2 is the EEG session and the only session that differs; the two non-EEG days
are statistically indistinguishable, differing by at most
`r round(d13_rng[2], 1)`% in any echo. The loss is present in **every** echo
(all 8 Day-2 contrasts in the figure survive Holm correction), and the magnitude
is stable: `r round(abs(pct_rng[2]), 1)`% to `r round(abs(pct_rng[1]), 1)`%
across the four echoes.

**The day x echo interaction is a scaling artefact.** In absolute units it is
significant
(*F*(`r round(gv(at,"day:echo","DFn"),2)`, `r round(gv(at,"day:echo","DFd"),2)`) =
`r round(gv(at,"day:echo","F"),2)`, *p* `r fmt_p(gv(at,"day:echo","p"))`,
ges = `r round(gv(at,"day:echo","ges"),3)`), but on the log scale it drops to
*F*(`r round(gv(atl,"day:echo","DFn"),2)`, `r round(gv(atl,"day:echo","DFd"),2)`) =
`r round(gv(atl,"day:echo","F"),2)`, *p* `r fmt_p(gv(atl,"day:echo","p"))`,
ges = `r signif(gv(atl,"day:echo","ges"),2)` — not significant. So the effect is
a constant *proportional* loss on Day 2, which looks like an interaction only
because echo 1 sits on a three-times-larger scale. That is the signature of a
multiplicative signal loss (or a noise gain) affecting all echoes equally,
consistent with the EEG cap and its leads rather than with anything TE-specific.

**The EEG penalty does not drift within the session.** Across the four Day 2
runs, mean tSNR is `r rm2m("Rest 2")` (rest2), `r rm2m("Rest 3")`,
`r rm2m("Rest 4")` and `r rm2m("Rest 5")` (rest5):
*F*(`r round(gv(at2,"run_f","DFn"),2)`, `r round(gv(at2,"run_f","DFd"),2)`) =
`r round(gv(at2,"run_f","F"),2)`, *p* `r fmt_p(gv(at2,"run_f","p"))`,
ges = `r round(gv(at2,"run_f","ges"),3)`, and the run x echo interaction is also
null (*p* `r fmt_p(gv(at2,"run_f:echo","p"))`,
ges = `r signif(gv(at2,"run_f:echo","ges"),2)`). `r n_sig_d2` of
`r nrow(run_by_echo)` within-echo run comparisons survive correction, which is
why the Day 2 figure carries no brackets. Rest3 is nominally the lowest run
(about `r round(abs(mean(pct_d2$"Rest 3 vs Rest 2 (%)")), 1)`% below rest2 in
every echo), but it does not survive correction and rest4 and rest5 return to
the rest2 level, so this is noise rather than a trend.

Taken together: the cost is attached to *having the cap on*, not to how long it
has been on. It appears at full size in the first EEG run and stays flat for the
rest of the session.

**Practical consequences.** Pooling resting-state runs across days is no longer
safe without accounting for session: a systematic
`r round(abs(d2_cost), 0)`% tSNR difference between the EEG day and the two
non-EEG days will propagate into any between-session comparison of effect sizes
or variance. Include session as a covariate, or restrict within-subject
contrasts to runs from the same day. Echo weighting for the multi-echo
combination, on the other hand, *can* use the same weights on each day — the
loss is proportional across echoes, so the relative weighting is unchanged.

## Does averaging tSNR over voxels make sense?

Partly. What is being analysed here is one number per subject, run and echo: the
mean of the voxelwise tSNR map inside that subject's grey-matter mask
(`fslstats -k … -M` in `eeg_fmri_ext_SNR_2_v2.sh`). For the question actually
being asked — *is image quality stable across sessions, and how does it scale
with echo?* — that is the right summary. It is a single scalar per cell, so the
repeated-measures ANOVA above is exactly the correct test, it has real power at
n = `r n_sub`, and restricting to grey matter already avoids the main artefact of
whole-brain averaging (ventricles and edge voxels dragging the mean around).

The averaging does, however, hide the things most likely to matter:

- **tSNR is strongly spatially non-uniform.** It is high in the cortex and low in
  orbitofrontal and inferior temporal regions, where susceptibility dropout is
  worst. A stable global mean is compatible with a real loss confined to those
  regions.
- **The global mean cannot localise the Day 2 loss.** A uniform
  `r round(abs(d2_cost), 0)`% drop in the GM mean is equally consistent with a
  uniform `r round(abs(d2_cost), 0)`% drop everywhere and with a much larger
  loss confined to electrode-adjacent cortex, offset by unchanged deep voxels.
  Those two possibilities have very different implications for whether task
  activation on Day 2 is comparable to the other days.
- **The mean is not where the optimal echo weighting lives.** Optimal T2\*
  weighting is voxelwise, since T2\* varies regionally; a global mean per echo
  cannot inform it.

## Does a voxel-wise analysis make sense?

Yes, and now it is the obvious next step rather than an optional complement.
Part 1 establishes *that* Day 2 costs tSNR; the maps already exist
(`*_bold_tSNR_masked.nii.gz`), so answering *where* is cheap. The sensible route
is to normalise each subject's tSNR maps to a common template, then fit the same
day x echo model voxelwise (SPM flexible factorial, or FSL randomise for a
non-parametric test), with cluster- or TFCE-based correction.

The prospects are much better than they were for a `r round(abs(d2_cost), 0)`%
effect that was borderline in the omnibus test. Here the day effect is
significant with ges = `r round(gv(at,"day","ges"),3)` and consistent across all
four echoes, so a voxelwise test is looking for something that is already known
to be present in the average — the question is only how it is distributed. Two
things still to plan for: tSNR maps are smooth and heavy-tailed, so a
non-parametric or log-transformed test is safer than assuming normality per
voxel; and because the effect is proportional, the voxelwise model should be fit
on `log(tSNR)` so that a uniform relative loss appears as a uniform additive
shift instead of tracking baseline tSNR.

The specific hypothesis worth testing is spatial: if the loss is caused by the
cap, it should be largest in superficial cortex near the electrodes and smallest
in deep and central structures. A small set of *a priori* ROIs — an
occipital/electrode-adjacent ROI, a dropout-prone orbitofrontal ROI, and a
deep/central ROI — analysed with the same day x echo model, tests that directly
at a fraction of the correction burden of a whole-brain search. A full voxelwise
map is then best used descriptively, to show the spatial distribution of the
Day 2 loss.

## Session info

```{r session}
sessionInfo()
```
