---
title: "tSNR across days and echoes"
author: "Benjamin Meyer"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 8, fig.height = 5.5, dpi = 200,
  dev.args = list(bg = "white")
)
library(ggplot2)
library(tidyr)
library(dplyr)
library(rstatix)
library(knitr)
```

## Read the data

The file is comma-separated: column 1 is the subject ID, the remaining columns
are tSNR values named `r<run>e<echo>` (R = resting-state scan, E = echo).
`NaN` entries (runs/echoes missing for a subject) are read as `NA`.

```{r read}
tsnr <- read.csv(
  "tSNR_results.txt",
  header     = TRUE,
  na.strings = c("NA", "NaN", "")
)

names(tsnr)[1] <- "ID"
value_cols <- setdiff(names(tsnr), "ID")

str(tsnr)
```

## Reshape to long format

`run` and `echo` are pulled out of the column names.

```{r reshape}
tsnr_long <- tsnr |>
  pivot_longer(
    cols      = all_of(value_cols),
    names_to  = "scan",
    values_to = "tSNR"
  ) |>
  mutate(
    scan = factor(scan, levels = value_cols),
    run  = as.integer(sub("^r(\\d+)e\\d+$", "\\1", scan)),
    echo = factor(as.integer(sub("^r\\d+e(\\d+)$", "\\1", scan)))
  ) |>
  filter(!is.na(tSNR))

head(tsnr_long)
```

## Restrict to runs 1, 2 and 6, relabelled as days

Runs 1, 2 and 6 are the first resting-state scan of each of the three
acquisition days. The mapping is taken from the acquisition loop in
`eeg_fmri_ext_SNR_2.sh` (`ses-Day1_task-rest1`, `ses-Day2_task-rest2`,
`ses-Day3_task-rest6`), and the EEG status from `eeg_fmri_ext_SNR_1.m`
("restscans 1, 6: without EEG; 2, 3, 4, 5: with EEG"):

| Run | Session | Label here | EEG |
|-----|---------|-----------|-----|
| 1 | `ses-Day1_task-rest1` | Day 1 | no |
| 2 | `ses-Day2_task-rest2` | Day 2 | **yes** |
| 6 | `ses-Day3_task-rest6` | Day 3 | no |

```{r subset}
runs_keep <- c(1, 2, 6)
day_labels <- c("Day 1", "Day 2", "Day 3")

tsnr_day <- tsnr_long |>
  filter(run %in% runs_keep) |>
  mutate(
    day  = factor(run, levels = runs_keep, labels = day_labels),
    scan = droplevels(scan),
    echo = droplevels(echo)
  )

# The ANOVA is a fully within-subject design, so it needs subjects with a value
# in every day x echo cell.
n_cells <- length(runs_keep) * nlevels(tsnr_day$echo)

tsnr_rm <- tsnr_day |>
  group_by(ID) |>
  filter(dplyr::n() == n_cells) |>
  ungroup() |>
  mutate(ID = factor(ID))

cat(
  "Complete cases: ", n_distinct(tsnr_rm$ID), " of ",
  n_distinct(tsnr_day$ID), " subjects\n", sep = ""
)
```

Subjects excluded, and which of the three runs they have:

```{r dropped}
tsnr_day |>
  group_by(ID) |>
  filter(dplyr::n() < n_cells) |>
  summarise(runs_present = paste(sort(unique(run)), collapse = ", "), .groups = "drop")
```

## Summary statistics

```{r summary-table}
tsnr_rm |>
  group_by(echo, day) |>
  summarise(
    n      = dplyr::n(),
    mean   = mean(tSNR),
    sd     = sd(tSNR),
    median = median(tSNR),
    IQR    = IQR(tSNR),
    min    = min(tSNR),
    max    = max(tSNR),
    .groups = "drop"
  ) |>
  mutate(across(mean:max, ~ round(.x, 2)))
```

## Plot

Black boxes and points, no background grid, explicit x- and y-axis lines, and
`base_size = 14` so the figure stays legible at A4 width. The x-axis is
two-tiered: echo number on the axis itself, day beneath.

```{r theme}
base_font <- 14   # readable when the 8 x 5.5 in figure is placed on A4

theme_tsnr <- theme_classic(base_size = base_font) +
  theme(
    panel.grid       = element_blank(),
    axis.line        = element_line(colour = "black", linewidth = 0.5),
    axis.ticks       = element_line(colour = "black", linewidth = 0.5),
    axis.text        = element_text(colour = "black"),
    axis.title       = element_text(colour = "black"),
    axis.title.x     = element_text(margin = margin(t = 30)),
    axis.title.y     = element_text(margin = margin(r = 8)),
    plot.title       = element_text(face = "bold", size = base_font + 1),
    plot.subtitle    = element_text(size = base_font - 2),
    plot.caption     = element_text(size = base_font - 3, hjust = 0, colour = "grey30"),
    plot.margin      = margin(6, 8, 14, 6)
  )

theme_set(theme_tsnr)
```

```{r plot, fig.width=8, fig.height=5.5}
n_echo <- nlevels(tsnr_rm$echo)

# One label per day, centred under that day's block of echoes.
day_tier <- data.frame(
  x     = seq_len(length(runs_keep)) * n_echo - (n_echo - 1) / 2,
  label = day_labels
)

# Jitter is horizontal only, so no point is drawn at a false tSNR value.
set.seed(42)

p_tsnr <- ggplot(tsnr_rm, aes(x = scan, y = tSNR)) +
  geom_boxplot(
    fill = "white", colour = "black",
    linewidth = 0.5, width = 0.62,
    outlier.shape = NA          # outliers are already shown as points below
  ) +
  geom_point(
    position = position_jitter(width = 0.16, height = 0, seed = 42),
    shape = 21, fill = "black", colour = "white",
    size = 2.1, stroke = 0.35, alpha = 0.9
  ) +
  geom_text(
    data = day_tier, aes(x = x, y = -Inf, label = label),
    vjust = 3.5, size = base_font / .pt, colour = "black",
    inherit.aes = FALSE
  ) +
  scale_x_discrete(labels = \(z) sub("^r\\d+e", "", z)) +
  scale_y_continuous(expand = expansion(mult = c(0.04, 0.06))) +
  coord_cartesian(clip = "off") +
  labs(
    title    = "Temporal SNR by echo and acquisition day",
    subtitle = "Each dot is one subject; Day 2 is the session acquired with EEG",
    x        = "Echo",
    y        = "tSNR (grey matter)",
    caption  = paste0("n = ", n_distinct(tsnr_rm$ID), " subjects with complete data")
  )

p_tsnr
```

```{r plot-save, include=FALSE}
# Vector copy at A4 text width, for inclusion in a manuscript.
ggsave("tSNR_boxplot.pdf", p_tsnr, width = 17, height = 12, units = "cm")
```

## Two-way repeated-measures ANOVA

Every subject contributes a value in all `r n_cells` cells, so **day** (3 levels)
and **echo** (4 levels) are both within-subject factors. One ANOVA, with both
main effects and their interaction. Greenhouse-Geisser is applied automatically
where Mauchly's test indicates a sphericity violation; `ges` is generalised
eta-squared.

```{r anova}
fit <- anova_test(tsnr_rm, dv = tSNR, wid = ID, within = c(day, echo))

get_anova_table(fit, correction = "auto")
```

Sphericity:

```{r sphericity}
fit$`Mauchly's Test for Sphericity`

fit$`Sphericity Corrections`
```

### Marginal means

```{r marginals}
tsnr_rm |>
  group_by(day) |>
  summarise(mean = round(mean(tSNR), 2), sd = round(sd(tSNR), 2), .groups = "drop")

tsnr_rm |>
  group_by(echo) |>
  summarise(mean = round(mean(tSNR), 2), sd = round(sd(tSNR), 2), .groups = "drop")
```

### Post-hoc comparisons

Paired *t*-tests between days, Holm-corrected. First collapsed across echoes
(follow-up to the day main effect), then within each echo (follow-up to the
interaction).

```{r posthoc}
pairwise_t_test(tsnr_rm, tSNR ~ day, paired = TRUE, p.adjust.method = "holm") |>
  select(group1, group2, statistic, df, p, p.adj, p.adj.signif)

tsnr_rm |>
  group_by(echo) |>
  pairwise_t_test(tSNR ~ day, paired = TRUE, p.adjust.method = "holm") |>
  select(echo, group1, group2, statistic, df, p, p.adj, p.adj.signif)
```

### Size of the day differences, per echo

The interaction is easier to read as a proportion, because tSNR itself differs
by a factor of three across echoes.

```{r pct-change}
tsnr_rm |>
  group_by(echo, day) |>
  summarise(mean = mean(tSNR), .groups = "drop") |>
  pivot_wider(names_from = day, values_from = mean) |>
  mutate(
    `Day 2 vs Day 1 (%)` = 100 * (`Day 2` - `Day 1`) / `Day 1`,
    `Day 3 vs Day 1 (%)` = 100 * (`Day 3` - `Day 1`) / `Day 1`,
    across(where(is.numeric), ~ round(.x, 2))
  )
```

### Is the interaction just a scaling effect?

An equal *proportional* change across echoes would still show up as an
interaction in absolute units, since echo 1 has three times the tSNR of echo 4.
Refitting on `log(tSNR)` turns proportional effects into additive ones, so if
the interaction were purely a scaling artefact it would disappear here.

```{r log-check}
tsnr_rm |>
  mutate(log_tSNR = log(tSNR)) |>
  anova_test(dv = log_tSNR, wid = ID, within = c(day, echo)) |>
  get_anova_table(correction = "auto")
```

## Summary of observed effects

```{r summary-numbers, include=FALSE}
at   <- get_anova_table(fit, correction = "auto")
gv   <- function(eff, col) at[[col]][at$Effect == eff]
n_sub <- n_distinct(tsnr_rm$ID)
```

**Echo dominates.** tSNR falls monotonically with echo number — `r round(mean(tsnr_rm$tSNR[tsnr_rm$echo == 1]), 1)`
(echo 1), `r round(mean(tsnr_rm$tSNR[tsnr_rm$echo == 2]), 1)`,
`r round(mean(tsnr_rm$tSNR[tsnr_rm$echo == 3]), 1)`, and
`r round(mean(tsnr_rm$tSNR[tsnr_rm$echo == 4]), 1)` (echo 4) — a very large
effect (*F*(`r round(gv("echo","DFn"),2)`, `r round(gv("echo","DFd"),2)`) =
`r round(gv("echo","F"),1)`, *p* < .001, ges = `r round(gv("echo","ges"),3)`).
This is the expected T2\* decay: later echoes have longer TE, so less signal
remains while thermal noise stays roughly constant. It is a property of the
sequence, not a data-quality finding.

**No effect of day.** Collapsed across echoes, mean tSNR is
`r round(mean(tsnr_rm$tSNR[tsnr_rm$day == "Day 1"]), 1)` (Day 1),
`r round(mean(tsnr_rm$tSNR[tsnr_rm$day == "Day 2"]), 1)` (Day 2) and
`r round(mean(tsnr_rm$tSNR[tsnr_rm$day == "Day 3"]), 1)` (Day 3):
*F*(`r round(gv("day","DFn"),2)`, `r round(gv("day","DFd"),2)`) =
`r round(gv("day","F"),2)`, *p* = `r round(gv("day","p"),3)`,
ges = `r round(gv("day","ges"),3)`. No pairwise comparison survives correction.
Image quality is stable across the three sessions, which is the reassuring
result for a multi-day design.

**A small day × echo interaction.**
*F*(`r round(gv("day:echo","DFn"),2)`, `r round(gv("day:echo","DFd"),2)`) =
`r round(gv("day:echo","F"),2)`, *p* = `r round(gv("day:echo","p"),3)`,
ges = `r round(gv("day:echo","ges"),3)`. It is carried by echo 1, which loses
about 3.6% of its tSNR on Day 2 (73.9 → 71.2) and recovers on Day 3 (74.5),
while echoes 2–4 vary by roughly 2% and do not dip. The interaction survives the
log transform, so it is not merely the consequence of echo 1 sitting on a larger
scale. Day 2 is the EEG session, and an EEG cap plausibly costs signal in the
shortest-TE echo first — but with *p* just under .05, ges = 0.007, and no
surviving post-hoc contrast, this should be treated as a hint to check rather
than an established effect. Runs 3–5 (also EEG, same day) are available and
would be the natural test of whether the echo-1 dip is reproducible.

Practically: the design is safe to pool across days, and echo weighting for the
multi-echo combination can use the same weights on each day.

### Does averaging tSNR over voxels make sense?

Partly. What is being analysed here is one number per subject, run and echo: the
mean of the voxelwise tSNR map inside that subject's grey-matter mask
(`fslstats -k … -M` in `eeg_fmri_ext_SNR_2.sh:104`). For the question actually
being asked — *is image quality stable across sessions, and how does it scale
with echo?* — that is the right summary. It is a single scalar per cell, so the
repeated-measures ANOVA above is exactly the correct test, it has real power at
n = `r n_sub`, and restricting to grey matter already avoids the main artefact of
whole-brain averaging (ventricles and edge voxels dragging the mean around).

The averaging does, however, hide the things most likely to matter:

- **tSNR is strongly spatially non-uniform.** It is high in the cortex and low in
  orbitofrontal and inferior temporal regions, where susceptibility dropout is
  worst. A stable global mean is compatible with a real loss confined to those
  regions.
- **Losses that trade off spatially cancel.** If the EEG cap costs tSNR near the
  electrodes (superficial, occipital and frontal cortex) while deeper voxels are
  unchanged, the global mean moves by very little — which may be exactly what the
  small echo-1 effect is.
- **The mean is not where the optimal echo weighting lives.** Optimal T2\*
  weighting is voxelwise, since T2\* varies regionally; a global mean per echo
  cannot inform it.

### Does a voxel-wise analysis make sense?

Yes, and it is worth doing as a complement rather than a replacement — the maps
already exist (`*_bold_tSNR_masked.nii.gz`), so the marginal cost is low. It
answers the question the global mean cannot: *where* does tSNR change. The
sensible route is to normalise each subject's tSNR maps to a common template,
then fit the same day × echo model voxelwise (SPM flexible factorial, or FSL
randomise for a non-parametric test), with cluster- or TFCE-based correction for
multiple comparisons.

Two things to plan for. First, the multiple-comparison penalty is real: a
whole-brain voxelwise test of the ~2% day effect seen here will very likely find
nothing, because the effect is small and diffuse rather than focal. Second, tSNR
maps are smooth and heavy-tailed, so a non-parametric or log-transformed test is
safer than assuming normality per voxel.

Given that, the most informative middle path is probably neither extreme: keep
the global GM mean as the omnibus quality check, and add a small set of
*a priori* ROI means — an occipital/electrode-adjacent ROI, a dropout-prone
orbitofrontal ROI, and a deep/central ROI — analysed with the same day × echo
model. That directly tests the EEG-proximity hypothesis suggested by the echo-1
dip, at a fraction of the correction burden of a whole-brain search. A full
voxelwise map is then best used descriptively, to show the spatial distribution
of tSNR per echo, rather than as the inferential test.

## Session info

```{r session}
sessionInfo()
```
